In Hops in order to improve performance and HA of the RM we have introduced an
in-memory distributed MySQL database which stores all the necessary
metadata. One great feature of HopsYarn is that the
\emph{ResourceTrackerService} (RT) of the RM is distributed into multiple
nodes in the cluster. That service is responsible for receiving and handling
heartbeats from the NMs. That way each instance handles only a portion of the total NM
heartbeats. The updated metadata are then stored into the database and
are streamed to the RM to update its view of the cluster. By load
balancing the ResourceTrackerService we have increased the performance
of the system while decreasing the load of the master RM which can
perform the rest of the operations without the load of handling every
single heartbeat.

Another equally important feature of HopsYarn is that RM stores
every event received and any scheduling decision into the MySQL
cluster. That makes our solution highly available with minimum
failover period. When a RM instance fails it re-builds the view of the
cluster by reading the latest state from the database. More details on
the architecture of both YARN and HopsYarn will be given in Chapter
\ref{chap:background} and \ref{chap:analysis}.

All these read/write operations to the database do not come without
a cost. First and foremost is the network latency. Even with high
throughput, low latency network in between of the RM/RT and the database, still
it takes more time than in-memory operations. Especially in cases
where RM operations need more than one round-trip to the database, the
difference in performance is noticeable.

A great advantage of using a relational database is the support of
foreign key constraints. The information that we store is semantically
related. In a SQL schema that is directly translated to foreign key
constraints. A trivial example is information regarding the containers
running on node and information regarding the node itself. Clearly, we
should never run into a situation where we try to persist a container
running on a node without having information about the node at
all. Moreover, foreign key constraints work the other way around. By
removing the information about a node from the database we want the
information about running containers on that node to be purged
too. Such constraints, although they seem tempting to use, pose a
great performance degradation. Particularly when we aim for millions
of operations per second, the use of foreign key constraints should be
limited and very well designed.