Large organizations and research institutes store a huge volume of data nowadays.
In order to gain any valuable insights distributed processing frameworks over a
cluster of computers are needed. Apache Hadoop is the prominent framework for distributed
storage and data processing. At SICS Swedish ICT we are building Hops, a new distribution
of Apache Hadoop supported by a distributed, highly available MySQL
Cluster NDB. Hops-YARN is the resource management framework of Hops
with distributed ResourceManagers load balancing the\\
ResourceTrackerService. In Hops-YARN we make heavy usage of the
back-end database storing all the ResourceManager's metadata and
incoming RPCs to provide high fault tolerance and very fast recovery
time. Also, the Event API of NDB is utilized as a
communication transport among the ResourceManager and the distributed
ResourceTrackers.

This project aims in optimizing the mechanisms used for persisting
metadata in NDB both in terms of transactional commit time but also
in terms of pre-processing them while at the same time guarantee the
consistency of RM's state. Under no condition should the in-memory RM
state diverge from the state stored in NDB. With these goals in mind
several solutions were examined that improved the performance of the
system, making Hops-YARN comparable to Apache YARN. The solutions
proposed in this thesis project enhance the pure commit time of a
transaction to the MySQL Cluster and the pre-processing and parallelism
of our Transaction Manager. The results indicate that the performance
of Hops increased dramatically, utilizing more resources on a cluster
with thousands of machines. Increasing the cluster utilization by a
few percentages can save organizations a big amount of money.
