In Hops-YARN, as it is already mentioned, MySQL Cluster NDB is used
both as a communication transport and as a persistent storage for
recovery. At the beginning of the project a thorough profiling of the
execution workflow has been done to identify the bottlenecks of the
system. The first step analyzed in Section \ref{sec:tx_aggregation}
was to improve the commit mechanism of Hops-YARN's Transaction
Manager. The new mechanism squashes several blocked transactions into
a ``big'' one, reducing the number of commits in the back-end database
system. Section \ref{sec:fk_constraints} describes the evolution to a
database schema with no foreign key constraints and how they were
replaced by application logic that performs primary key operations instead.
Section \ref{sec:gc_service} presents the Garbage Collector
service of Hops-YARN that asynchronously removes old values from the
database. With that solution the commit time dropped even more improving
the overall performance of the system. Finally, Section
\ref{sec:dto_caching} explains how the shortcoming of ClusterJ for
creating DTOs was bypassed by having created them ahead of time in a
per session cache.

After a detailed explanation of the solutions proposed, in Chapter
\ref{chap:evaluation} follows the evaluation. Each solution is
evaluated separately by simulating real world traces. In each case key
characteristics are examined and how they have been improved. In
Section \ref{sec:performance_overview} there is an overall performance
overview in terms of cluster utilization and heartbeats processed by
the scheduler. The comparison is made among the version of Hops-YARN
before this project, the final version with all the modification
proposed and the upstream Apache YARN. The figures show that there was
a clear improvement, in both evaluation parameters, when the two
version of Hops-YARN are compared. Finally, as far as the cluster
utilization is concerned, the performance is comparable with Apache
YARN in clusters with thousands of nodes.